Comparison of sediment transport capacity variations in subglacial and subaerial systems

### 2023-03-07 theory.tex
polishing

\newcommand{\aref}[1]{\textbf{Reference #1}}
\newcommand{\TODO}[1]{\textbf{TODO #1}}
\newcommand{\ian}[1]{{\textbf{\color{blue}Ian says:} \color{blue} #1} }
\newcommand{\m}{$\,\mathrm{m}$\,}
\newcommand{\cm}{$\,\mathrm{cm}$\,}
\newcommand{\mma}{$\,\mathrm{mm  \, a^{-1}}$\,}
\newcommand{\mmma}{$\,\mathrm{m^3\, a^{-1}}$\,}
\newcommand{\unit}[1]{$\mathrm{#1}$}

@@

\newenvironment{abstract_}{
  \chapter{\centering \textbf{Abstract}}
}

\author[1]{Ian Delaney}
\affil[1]{Institut des dynamiques de la surface terrestre (IDYST), Universit\'{e} de Lausanne, B\^{a}timent G\'{e}opolis, CH-1015 Lausanne}
\title{Sediment transport capacity varies more in subglacial systems compared to fluvial ones}

@@

\paragraph{Key findings:}
\begin{enumerate}
\item Sediment transport capacity varies more in pressurized subglacial channels compared to open fluvial ones, in response to changing water discharge due to fixed conduit size of their channels.
\item Differences in the sediment transport capacity response to changing water discharge could impact processes at proglacial margins and the interpretation of sediment transport records from glaciers. 
\end{enumerate}

\abstract % 150 words
Sediment transport capacity depends largely on the velocity of water flowing through a channel and the channel width over which to mobilize sediment.
In fluvial channels, changing water discharge can be accommodated by water depth, width, and water velocity.
In subglacial channels, over short time scales, however, water flows through a conduit with a fixed size, and the glacier ice above pressurizes the water.
As a result, changing water discharge only may result in modifying water velocity, not the channel's shape.
Parameterizations of flow in both pressurized, subglacial channels and  open, fluvial channels show that sediment transport capacity varies  more in response to changing water discharge in subglacial channels compared to fluvial ones.
This characteristic of increased variability in sediment transport capacity in subglacial systems holds true  across a wide range of channel geometries and flow conditions.
The increased variability in sediment transport capacity impacts sediment transport processes and the interpretation of records of sediment transport from glacierized catchments.

@@

\laysummary % 200 words
Sediment transport capacity responds differently to changing water discharge in subglacial and fluvial channels.
In both subglacial and fluvial channels, sediment transport capacity, to the first order, depends on 1) the channel width over-which to mobilize sediment and 2) the water's shear stress on sediment at the channel bottom to mobilize it.
The shear stress depends largely on water velocity.
In fluvial channels, changing water discharge will change the water depth, water velocity (controlling sediment transport capacity),  and channel width (which dictates the span over which sediment transport  will occur).
Conversely, in subglacial environments, on short times scales of hours to days, subglacial channels only experience a change in water velocity following changing water discharge, because the water flows through a subglacial conduit with a fixed geometry.
The fixed conduit geometry also means that the channel width, over-which sediment can be mobilized, does not respond to changing water discharge.
As a result, sediment transport capacity varies far more in subglacial channels compared to fluvial ones, even though changing channel width can accommodate some sediment transport variability in fluvial channels.
This increased variability has several important implications for sediment transport in glacial systems compared to fluvial ones.

@@

  Glaciers are well known for the quantities of sediment that they expel from their termini \citep{hallet1996}.
  Changing glacier dynamics, geomorphology, and hydrology  has prompted a recent focus on  sediment transport processes in cold regions \citep[e.g.][]{zhang2022}.          
  Numerical models consider the subglacial component of sediment discharge as a function of the availability of sediment, from the erosion of bedrock or the prior existence of material, and the sediment transport capacity by fluvial processes below the glacier \citep{creyts2013,brinkerhoff2017,beaud2018,delaney2019}.
  These models often represent sediment transport processes similar to rivers, with the main difference being that water flows through a subglacial conduit, pressurized by the ice above \citep{rothlisberger1972}.

  Both subglacially and fluvially, shear stress between the water and the sediment that it flows over is a first-order control on the transport of sediment \citep{shields1936} if an adequate quantity of sediment is present as to be transported fluvially (i.e. in a ``transport-limited'' regime). 
  Shear stress increases with the water velocity and roughness of the channel.
  Greater shear stress will mobilize increased quantities of sediment, along with larger sediment clasts.
  Sediment transport relationships such as \citet{shields1936}, \citet{meyer1948},  and \citet{engelund1967} use shear stress, or an approximation thereof, to evaluate sediment transport capacity per unit width.
  The total sediment transport capacity across a channel's cross depends on the integrated shear stress over the entire channel bed (Figure~\ref{fig:cartoon}).
  As a result, channel geometry plays a key role in sediment transport in both fluvial channels and subglacial ones \citep[e.g.][]{church2006,delaney2019}.

  However, in response to changing water discharge, the behavior of subglacial channels differs from fluvial ones.
  In subglacial channels, the water is pressurized in a subglacial channel and the size of this channel responds to the roughness of the channel walls and the velocity of the water flowing through the channel \citep[Figure~\ref{fig:cartoon}\,a and c; e.g. ][]{rothlisberger1972}.
  Here, diurnal variations in water flux, for instance, may cause water to flow through a conduit whose size does not respond to the same timescales as the increase in water flux \citep{rothlisberger1972}.
  With respect to sediment transport, this means that increased water discharge in subglacial channels will result in greater water velocity and thus sediment transport capacity.
  At the same time, the fixed channel width in subglacial channels means that the channel width over which to mobilize sediment does not evolve over the same timescales as water discharge \citep[e.g.][]{werder2010}.

  Conversely, in fluvial systems with open-channel flow, a commensurate increase in water discharge will be accommodated by both a reduced increase in water velocity, along with increased channel width and/or depth (Figure~\ref{fig:cartoon}\,b and d).
  The change in water velocity could be lower, but the width over which sediment can be mobilized changes \citep{leopold1953}.
  Thus, the systematic response in sediment transport capacity with respect to changing water discharge is  different in the two systems.

  The different processes in subglacial and fluvial systems have been implicitly included in a wide range of available models and relationships that transport sediment in both glacial and fluvial environments \citep[e.g.][]{walder1994,tucker1997,beaud2018,wickert2019}.
  Furthermore, models of fluvial transport of sediment in glacial and fluvial systems have been used to understand the relationship between sediment dynamics and climate \citep[e.g.][]{tucker1997,delaney2020}
  Yet, the differing relationship between variations in water discharge and sediment discharge in subglacial and fluvial systems has been less well discussed.

  To compare the differences in the response of sediment transport capacity to water discharge, this manuscript leverages parameterizations of subglacial hydraulics and fluvial hydraulics.
  Parameterizations are applied to explore the differences in the relationship between sediment transport capacity and water discharge and the impact of water discharge variability on sediment transport capacity in the two systems.
  Lastly, the implications for sediment transport processes in subglacial systems, compared to fluvial ones, are discussed.

  \begin{center}
    \begin{figure}[H]
      \includegraphics[width=0.65\linewidth]{Cartoon.pdf}
      \caption{Cartoon of the situation with various parameters. Velocity $v_g$ and $v_f$ magnitude are shown by arrow length and marks at the channel boundary. $S_g$ and $S_f$ represent the cross-sectional area in glacial and fluvial channels.  Note that in the fluvial channel parameterization, rectangular channel shapes are implemented (Section~\ref{sect:fluv}).} 
      \label{fig:cartoon}
    \end{figure}
  \end{center}

  \section{Methodology}
  \label{sect:meth}
  This section contains the relationships between water discharge, subglacial water velocity, and channel geometry in both fluvial and subglacial channels.
  Because sediment transport largely depends on shear stress \citep{shields1936}, both parameterizations calculated this value and integrate  it across the channel bed.
  The parameterizations evaluate shear stress, as opposed to sediment transport capacity, as shear stress omits a grain size parameter and the choice of a sediment transport relationship \citep{shields1936}.

  \ian{check units on fluvial channel factor}
  \begin{table}[H]
    \centering
    \caption{Variables, parameters, and constants }
    \begin{tabular}{ l  c  c c }
      Name &Symbol&  Value&Units \\ \hline
      \textbf{Variables}  & & & \\
      Channel hydraulic diameter (glacier, fluvial) &  $D_{hg},\,D_{hf}$&  & $\mathrm{m}$     \\
      Width of channel floor (glacial, fluvial) & $w_{cg},w_{cf}$&  & $\mathrm{m}$     \\
      Channel cross-sectional area (glacier, fluvial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
      Water discharge (instantaneous) & $Q_w$& & $\mathrm{m^{3}\,s^{-1}}$ \\
      Representative water discharge & $Q_{w}^*$& & $\mathrm{m^{3}\,s^{-1}}$ \\
      % Gradient of hydraulic potential (glacier) &$\Psi$ & & $\mathrm{Pa\, m^{-1}}$\\
      Representative hydraulic gradient  &$\Psi^*$ & & $\mathrm{Pa\, m^{-1}}$\\
      Water velocity (glacier, fluvial)  & $v_g,\,v_{f}$& & $\mathrm{m\,s^{-1}}$ \\
      Shear stress (glacier, fluvial) & $\tau_g,\,\tau_f$&& $\mathrm{Pa \, m^{-1}}$ \\
      Width-integrated shear-stress (glacier, fluvial) & $\tau_{tg},\, \tau_{tf}$&& $\mathrm{Pa \, m^{-1}}$ \\
      Reynolds number &$R_e$& & $\mathrm{(-)}$\\
      Variable &$\Gamma$&&$\mathrm{(-)}$\\
      Variability of a variable ($\Gamma$) &$\chi$& &$\mathrm{(-)}$\\
           &&&\\

      \textbf{Parameters and Constants}  & & &\\
      Gravitational constant&$g$& $-9.81$&$\mathrm{m\,s^{-2}}$\\
      Density of water & $\rho_w$& $1000$ & $\mathrm{kg\,m^{-3}}$ \\
      Density of ice & $\rho_i$& $900$ & $\mathrm{kg\,m^{-3}}$ \\
      Kinematic viscosity of water &$\nu$& $10^{-4}$& $\mathrm{m^2\,s^{-1}}$\\
      Hooke angle of channel & $\beta$ & $\frac{\pi}{2}$ & \unit{rad}\\
      Flotation fraction & $f_f$&$0.7$& $\mathrm{(-)}$\\
      Friction factor (glacier) & $f_r$ & $4$ & $\mathrm{(-)}$ \\
      Friction factor (fluvial) & $f_p$ & $3$ & $\mathrm{(-)}$\\
      Gradient of glacier surface & $\frac{\partial z_{sg}}{\partial x}$ &$0.25$& $\mathrm{(-)}$\\
      Gradient of channel bed (fluvial) &$\frac{\partial z_c}{\partial x}$ &$0.05$& $\mathrm{(-)}$\\
      Fluvial channel factor & $k$ &$3$ & $\mathrm{m^{-2}\, s}$\\
      Channel geometry exponent &$e$& $\frac{1}{2}$&$\mathrm{(-)}$ \\
      \hline
    \end{tabular}
    \label{table:vpm}
  \end{table}

  \subsection{Subglacial channel  parameterization}
  \label{sect:sub_mode}
  To parameterize the shear stress of water flowing across sediments below a glacier, the subglacial channel parameterization evaluates the channel geometry and the velocity of the flowing water. 
  To accomplish this, the hydraulics parameterization presented in \citet{delaney2019} is used.
  The parameterization uses the assumption that the water is transported through subglacial channels \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972}, and that a channel and geometry will respond to a representative water discharge over a certain time period $Q_{w}^*$, using the Darcy-Weisbach formulation for water-flow through a pipe to describe this relationship \citep[e.g.][]{rothlisberger1972,clarke2003,werder2013}.
  Here, it is assumed that the hydraulic diameter of the subglacial channel, $D_h$, responds to a representative water discharge, $Q_{w}^*$, and a representative gradient of the hydraulic potential, $\Psi^*$:
  \begin{linenomath*}
    \begin{equation}
      \label{eq:DW}
      D_h = \big(s\, f_r\,\rho_w\, \frac{Q_w^{*\,2}}{\Psi^*}\big)^{\frac{1}{5}}~~.
    \end{equation}
  \end{linenomath*}
  % 
  The representative hydraulic gradient is based upon the pressurized flow in Equation~\ref{eq:DW},
  \begin{equation}
    \label{eq:psi}
    \Psi^*= f_f \,  \rho_i \, g (\frac{\partial  z_{sg}}{\partial x} - \frac{\partial z_c}{\partial x}) +  \rho_i \, g \, \frac{\partial z_c}{\partial x},
  \end{equation}
  % 
  \noindent
  where, $f_f$ is the flotation fraction, $\rho_i$ is density of ice $\frac{\partial z_{sg}}{\partial x}$ is the surface slope of the glacier and $\frac{\partial z_c}{\partial x}$ is the glacier bed slope.

  In Equation~\ref{eq:DW}, $f_r$ is the Darcy-Weisbach friction factor and $\rho_w$ is the density of water (Table \ref{table:vpm}). $s$ is a factor accounting for channel geometry \citep{hooke1990}, calculated as:
  \begin{equation}
    \label{eq:Hf}
    s = \frac{2\,(\beta -\sin \beta)^2}{(\frac{\beta}{2}\,+\,\sin \frac{\beta}{2})^4},
  \end{equation}
  where $\beta$ is the central angle of the circular segment that comprises the channel (the so-called Hooke angle).
  Smaller values of $\beta$ result in shallow, wide channels. $\beta =\pi$ corresponds to a semi-circle.
  The width of the flat channel floor $w_c$ is given by
  \begin{equation}
    \label{eq:dh2wc}
    w_{cg} = 2  \sin \frac{\beta}{2} \sqrt{\frac{2\, S}{\beta -\sin \beta}},
  \end{equation}
  % 
  where $S$ is the cross-sectional area of the channel given by
  \begin{equation}
    \label{eq:dh2S}
    S_g =  \frac{D_h^2}{2}~ \frac{\Big(\frac{\beta}{2} \,+ \, \sin \frac{\beta}{2}\Big)^2  }{\beta\,-\,\sin \beta}.
  \end{equation}

  The shear stress, $\tau$, between the water and the channel bed is determined through the Darcy-Weisbach formulation
  \begin{equation}
    \label{eq:tau}
    \tau_g=\frac{1}{8}\,f_r\,\rho_w\,v_g^2,
  \end{equation}
  % 
  where $v_g = \frac{Q_w}{S_g}$ is the water velocity.
  Here, $Q_w$ represents the instantaneous water discharge, not used to evaluate the hydraulic gradient of the channel in Equation~\ref{eq:DW}.
  The channel size is usually not in equilibrium with the quantity of water flowing through it.
  With this formulation, the processes in an R-channel can be represented algebraically \citep{delaney2019}.

  For purposes discussed below, we evaluate the width integrated shear stress as $\tau_{tg}=w_{cg}\,\tau_g $.

  \subsection{Fluvial channel  parameterization}
  \label{sect:fluv}

  The fluvial channel parameterization implements the hydraulics parameterization presented in \citet{tucker1997}.
  Here, the assuming the conservation of mass, sufficiently wide channels so that the hydraulic radius is consistent with flow depth, uniform flow, and the Darcy-Weisbach relationship,
  the shear stress $\tau_f$ at the river bed is represented as
  \begin{linenomath*}
    \begin{equation}
      \label{eq:DW_tau}
      \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_p^{\frac{1}{3}}}{2}\, \Big(\frac{Q_w}{w_{cf}} \Big)^{\frac{2}{3}} \,\frac{\partial z_c}{\partial x}^{\frac{2}{3}},
    \end{equation}
  \end{linenomath*}
  where $\frac{\partial z_c}{\partial x}$ is the channel slope, and $f_p$ is the friction factor for fluvial systems.
  Channel width $w_{cf}$ is 
  \begin{equation}
    \label{eq:wcf}
    w_{cf} = k \, Q_w^e,
  \end{equation}
  % 
  where $k$ is a constant and $e$ is an exponent commonly equal to $\frac{1}{2}$ \citep{leopold1953}. Cross-sectional area is 
  \begin{equation}
    \label{eq:Sf}
    S_f= \frac{Q_w}{v_f}.
  \end{equation}
  As in Section~\ref{sect:sub_mode}, the width integrated shear stress is $\tau_{tf}=w_{cf}\,\tau_f$.
  Hydraulic diameter of the channel is given by $D_{hf} = \frac{4\,w_{cf}\,d}{2\,d+w_{cf}}$, with flow depth $d$ determined from knowledge of $S_f$ and $w_{cf}$ in a rectangular channel.

@@

  The formulation for the glacial system above requires inputs of the hydraulic gradient, water discharge, representative water discharge, Hooke angle $\beta$ \citep{hooke1990}, and friction factor $f_r$.
  Values of $\beta$ and $f_r$ are chosen so that they produce reasonable values of water velocity, in accordance with available measurements through dye-tracing \citep[Section~\ref{sect:sub_mode}, Figure~\ref{fig:model_outs}; e.g.][]{werder2010}.

  In the fluvial system, the water is not pressurized, so only the bed slope is needed. Friction factor $f_f$ and channel shape factor $k$ are assigned (Section~\ref{sect:fluv}).

>>>>

\newcommand{\aref}[1]{\textbf{Reference #1}}
\newcommand{\TODO}[1]{\textbf{TODO #1}}
\newcommand{\ian}[1]{{\textbf{\color{blue}Ian says:} \color{blue} #1} }
\newcommand{\marie}[1]{{\textbf{\color{green}Marie says:} \color{green} #1} }
\newcommand{\m}{$\,\mathrm{m}$\,}
\newcommand{\cm}{$\,\mathrm{cm}$\,}
\newcommand{\mma}{$\,\mathrm{mm  \, a^{-1}}$\,}
\newcommand{\mmma}{$\,\mathrm{m^3\, a^{-1}}$\,}
\newcommand{\unit}[1]{$\mathrm{#1}$}

@@

\newenvironment{abstract_}{
  \chapter{\cemuchntering \textbf{Abstract}}
}

\author[1]{Ian Delaney}
\affil[1]{Institut des dynamiques de la surface terrestre (IDYST), Universit\'{e} de Lausanne

 B\^{a}timent G\'{e}opolis, CH-1015 Lausanne}
\title{Comparison of sediment transport capacity variations in subglacial and subaerial systems }

@@

\paragraph{Key findings:}
\begin{enumerate}
\item The sediment carrying capacity of subglacial water varies more in response to changing water discharge compared to subaerial channels, as subglacial channels are pressurized by the ice above.
\item The slowly evolving size of subglacial channels means that variations in water discharge must be accommodated by changing the velocity of water flowing through the channel.
%\marie{I can't fully understand this. Too much information is overlain here. Separate concepts and give a more linear explanation.} 
\item  As water discharge changes, differences in the sediment transport capacity response of subaerial vs subglacial channels  could substantially  impact sediment transport processes at proglacial margins as well as the interpretation of sediment transport records in glacial environments. 
\end{enumerate}

\abstract % 150 words
Sediment transport capacity depends mainly on the velocity of water flowing through a channel and the channel width over which the sediment is mobilized.
In subaerial channels, changing water discharge can be accommodated by changes in the water depth, the width of the channel that  water flows over, and water velocity.
In subglacial channels, however, water flows through a conduit with a fixed size and shape, and the water is pressurized by the overlying glacier ice.
As a result, changing water discharge may modify the water velocity, but not the  shape of the channel.
Parameterization of channel dynamics indicates that increased variability in sediment transport capacity in subglacial systems occurs across a wide range of channel geometries and flow conditions.
The increased variability in sediment transport capacity impacts sediment transport processes and the interpretation of records of sediment transport from glacierized catchments.  %\marie{This last sentence seems repetitive. I do not fully understand the results. Suggest revision after a complete update of the manuscript}

@@

\laysummary % 200 words
It is well known that sediment transport capacity responds differently to changing water discharge in subglacial and subaerial channels.
In both subglacial and subaerial channels, sediment transport capacity depends, in the first order, on 1) the channel width over-which the sediment is mobilized  and 2) the shear stress applied by flowing water on sediment at the channel bottom.
The shear stress depends largely on water velocity.
In subaerial channels, changing water discharge will change the water depth, the water velocity (controlling sediment transport capacity),  and channel width (which dictates the span over which sediment transport  will occur).
Conversely, in subglacial environments, on short times scales of hours to days, subglacial channels only experience a change in water velocity following changing water discharge, because the water flows through a subglacial conduit with a fixed geometry.
The fixed conduit geometry also means that the channel width, over-which sediment can be mobilized, does not respond to changing water discharge.
As a result, sediment transport capacity varies far more in subglacial channels compared to subaerial ones, even though changing channel width can accommodate some sediment transport variability in subaerial channels.
This increased variability has several important implications for sediment transport in glacial systems compared to subaerial ones.
%\maire{This is hard to follow. Suggest you rewrite after greater clarity in the rest of the paper.}

@@

  Glaciers expel large quantities of sediment from their termini \citep{hallet1996}.
  Changes in glacier dynamics, geomorphology, and hydrology  have prompted numerous  recent studies of  sediment transport processes in cold regions \citep[e.g.][]{zhang2022}.
  Increased sediment discharge at both the glacier and regional scale \citep{costa2017,delaney2020,li2022,vergara2022}, creates an imperative to examine the processes controlling sediment discharge from glacierized catchments.

Over long periods, processes such as glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from below glaciers \citep[c.f.][]{hallet1979,iverson1993,herman2021}. 
 On shorter time periods, pressurized water transports available sediment from the glacier \citep{creyts2013,beaud2018b,delaney2019}, should enough sediment be present below the glacier (i.e. in a "transport"-limited regime).

In a "transport"-limited regime, sediment discharge responds to the sediment transport capacity or the amount of sediment that could be carried by the water. Sediment transport capacity depends on the shear stress between water and sediment it flows  over \citep{shields1936,meyer1948,engelund1967} and the width of the channel $w$ over which to mobilize sediment. The shear stress  across the channel width $\tau_t$ responds to the velocity of water $v$ flowing through at channel so that 
\begin{linenomath*}
    \begin{equation}
      \label{eq:tau_t}
        \tau_t \propto w\, v^2,
        \end{equation}
  \end{linenomath*}
  % 
  where $w$ is the width of the channel.
  Following the conservation of mass, the velocity of the water flowing through a rectangular channel is given as 
  \begin{linenomath*}
    \begin{equation}
      \label{eq:v}
  v = \frac{Q_w}{S(w)},
  \end{equation}
  \end{linenomath*}
%
where $S$ is the channel's cross-sectional area that responds to channel width $w$, and $Q_w$ is water discharge.

In subaerial channels, such as rivers with open channel flow, the depth $d$ and width $w$ of the channel evolve with changing water discharge $Q_w$ \citep{leopold1953}. During a flood with greater water discharge $Q_w$, for instance, cross-sectional area $S$ increases with channel width $w$ along with velocity $v$. As a result, the shear stress across the channel $\tau_t$ will increase in response to both width $w$ and $v$.

Yet, the response of water velocity to changing water discharge in subglacial channels is different than subaerial ones.
In subglacial systems, the water is pressurized by the ice above \citep{shreve1972}.
The size of subglacial channels responds to the creep closure of the ice above and the opening of the channel by frictional heating of water flowing through the channel \citep{rothlisberger1972}. As a result, the subglacial channel size can evolve on a longer timescale than variations in water discharge \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
Changes in water discharge $Q_w$ are mainly accommodated in water velocity $v$, because the width $w$ and depth $d$ of the channel are fixed (Figure~\ref{fig:cartoon}).
It follows that greater water discharge $Q_w$ will cause the shear stress across the channel $\tau_t$ to primarily respond to the water velocity $v$, as opposed to the subaerial system where width $w$ may evolve as well.

As a result, sediment mobilization in subaerial and subglacial systems responds differently to changing water discharge.
These differences have been implicitly included in a wide range of available models that  quantify sediment transport from these environments in both subglacial and subaerial systems  \citep[e.g.][]{walder1994,tucker1997,creyts2013,wickert2019,delaney2019}.
  The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \citep[e.g.][]{lane2016}.
  Additionally, the variable response of sediment transport capacity to water discharge in the two systems may impact the interpretation of records of sediment transport in glacier compared to subaerial systems \citep[e.g.][]{muller1968,richards2003}.
  Yet, the differing relationship between variations in water discharge and sediment discharge in subglacial and subaerial systems has been minimally discussed.

 This manuscript leverages parameterizations of subglacial and subaerial hydraulics to explore differences in the relationships between  water discharge, channel geometry, and water velocity. 
 With these parameterizations, the manuscript discusses the impact of water discharge variability on sediment transport capacity in idealized subglacial and subaerial systems.
  Lastly, the implications for sediment transport processes in subglacial systems, compared to subaerial ones, are discussed.

  \begin{center}
    \begin{figure}[H]
      \includegraphics[width=0.65\linewidth]{Cartoon.pdf}
      \caption{Cartoon of the situation with various parameters. Water velocity magnitude in the subglacial, $v_g$ , and subaerial ,$v_f$ , systems are shown by arrow length \marie{and lines of xxx greatest value} marks at the channel boundary. $S_g$ and $S_f$ represent the cross-sectional area in glacial and subaerial channels. Note that in the subaerial channel parameterization, rectangular channel shapes are implemented (Section~\ref{sect:fluv}).} 
      \label{fig:cartoon}
    \end{figure}
  \end{center}

  \section{Methods}
  \label{sect:meth}
 Relationships among water discharge, subglacial water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}) are described with subglacial and subaerial parameterizations.
  Because sediment transport largely depends on shear stress applied by water flow \citep{shields1936}, both parameterizations calculate this value and integrate  it across the channel bed (Fig. \ref{fig:cartoon}).
  The evaluation of shear stress  omits a grain size parameter and, thus, leaves open the selection of a sediment transport relationship  \citep{shields1936} and thus the calculation of sediment transport capacity.

  \ian{check units on subaerial channel factor}
  \begin{table}[H]
    \centering
    \caption{Variables, parameters, and constants }
    \begin{tabular}{ l  c  c c }
      Name &Symbol&  Value&Units \\ \hline
      \textbf{Variables}  & & & \\
      Channel hydraulic diameter (glacier, subaerial) &  $D_{hg},\,D_{hf}$&  & $\mathrm{m}$     \\
      Width of channel floor (glacial, subaerial) & $w_{cg},w_{cf}$&  & $\mathrm{m}$     \\
      Channel cross-sectional area (glacier, subaerial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
      Water discharge (instantaneous) & $Q_w$& & $\mathrm{m^{3}\,s^{-1}}$ \\
      Representative water discharge & $Q_{w}^*$& & $\mathrm{m^{3}\,s^{-1}}$ \\
      % Gradient of hydraulic potential (glacier) &$\Psi$ & & $\mathrm{Pa\, m^{-1}}$\\
      Representative hydraulic gradient  &$\Psi^*$ & & $\mathrm{Pa\, m^{-1}}$\\
      Water velocity (glacier, subaerial)  & $v_g,\,v_{f}$& & $\mathrm{m\,s^{-1}}$ \\
      Shear stress (glacier, subaerial) & $\tau_g,\,\tau_f$&& $\mathrm{Pa \, m^{-1}}$ \\
      Width-integrated shear-stress (glacier, subaerial) & $\tau_{tg},\, \tau_{tf}$&& $\mathrm{Pa \, m^{-1}}$ \\
      Reynolds number &$R_e$& & $\mathrm{(-)}$\\
      Variable &$\Gamma$&&$\mathrm{(-)}$\\
      Variability of a variable ($\Gamma$) &$\chi$& &$\mathrm{(-)}$\\
           &&&\\

      \textbf{Parameters and Constants}  & & &\\
      Gravitational constant&$g$& $-9.81$&$\mathrm{m\,s^{-2}}$\\
      Density of water & $\rho_w$& $1000$ & $\mathrm{kg\,m^{-3}}$ \\
      Density of ice & $\rho_i$& $900$ & $\mathrm{kg\,m^{-3}}$ \\
      Kinematic viscosity of water &$\nu$& $10^{-4}$& $\mathrm{m^2\,s^{-1}}$\\
      Hooke angle of channel & $\beta$ & $\frac{\pi}{2}$ & \unit{rad}\\
      Flotation fraction & $f_f$&$0.7$& $\mathrm{(-)}$\\
      Friction factor (glacier) & $f_r$ & $4$ & $\mathrm{(-)}$ \\
      Friction factor (subaerial) & $f_p$ & $3$ & $\mathrm{(-)}$\\
      Gradient of glacier surface & $\frac{\partial z_{sg}}{\partial x}$ &$0.25$& $\mathrm{(-)}$\\
      Gradient of channel bed (subaerial) &$\frac{\partial z_c}{\partial x}$ &$0.05$& $\mathrm{(-)}$\\
      Subaerial channel factor & $k$ &$3$ & $\mathrm{m^{-2}\, s}$\\
      Channel geometry exponent &$e$& $\frac{1}{2}$&$\mathrm{(-)}$ \\
      \hline
    \end{tabular}
    \label{table:vpm}
  \end{table}

  \subsection{Subglacial channel  parameterization}
  \label{sect:sub_mode}
  To parameterize the shear stress of water flowing across sediments below a glacier, the subglacial channel parameterization evaluates the channel geometry and the velocity of the flowing water. 
  To accomplish this, the hydraulics parameterization presented in \citet{delaney2019} is used. Here, it is assumed that the water is transported through subglacial channels \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972}, and that the channel  and geometry will respond to a representative water discharge over a certain time period $Q_{w}^*$, using the Darcy-Weisbach formulation for water-flow through a pipe  \citep[e.g.][]{rothlisberger1972,clarke2003,werder2013}.
  Thus, the hydraulic diameter of the subglacial channel, $D_h$, responds to a representative water discharge, $Q_{w}^*$, and a representative gradient of the hydraulic potential, $\Psi^*$ such that:
  \begin{linenomath*}
    \begin{equation}
      \label{eq:DW}
      D_h = \big(s\, f_r\,\rho_w\, \frac{Q_w^{*\,2}}{\Psi^*}\big)^{\frac{1}{5}}~~.
    \end{equation}
  \end{linenomath*}
  % 
  The representative hydraulic gradient is based upon the pressurized flow in Equation~\ref{eq:DW},
  \begin{equation}
    \label{eq:psi}
    \Psi^*= f_f \,  \rho_i \, g (\frac{\partial  z_{sg}}{\partial x} - \frac{\partial z_c}{\partial x}) +  \rho_i \, g \, \frac{\partial z_c}{\partial x},
  \end{equation}
  % 
  \noindent
  where, $f_f$ is the flotation fraction, $\rho_i$ is density of ice $\frac{\partial z_{sg}}{\partial x}$ is the surface slope of the glacier and $\frac{\partial z_c}{\partial x}$ is the glacier bed slope.

  In Equation~\ref{eq:DW}, $f_r$ is the Darcy-Weisbach friction factor and $\rho_w$ is the density of water (Table \ref{table:vpm}). $s$ is a factor accounting for channel geometry \citep{hooke1990}, calculated as:
  \begin{equation}
    \label{eq:Hf}
    s = \frac{2\,(\beta -\sin \beta)^2}{(\frac{\beta}{2}\,+\,\sin \frac{\beta}{2})^4},
  \end{equation}
  where $\beta$ is the central angle of the circular segment that comprises the channel (the so-called Hooke angle). Note that $\beta =\pi$ corresponds to a semi-circle and
  smaller values of $\beta$ result in shallow, wide channels. 
  The width of the flat channel floor $w_c$ , given $beta$ is   \begin{equation}
    \label{eq:dh2wc}
    w_{cg} = 2  \sin \frac{\beta}{2} \sqrt{\frac{2\, S}{\beta -\sin \beta}},
  \end{equation}
  % 
  where $S$ is the cross-sectional area of the channel given by
  \begin{equation}
    \label{eq:dh2S}
    S_g =  \frac{D_h^2}{2}~ \frac{\Big(\frac{\beta}{2} \,+ \, \sin \frac{\beta}{2}\Big)^2  }{\beta\,-\,\sin \beta}.
  \end{equation}

  The shear stress, $\tau$, between the water and the channel bed is determined through the Darcy-Weisbach formulation
  \begin{equation}
    \label{eq:tau}
    \tau_g=\frac{1}{8}\,f_r\,\rho_w\,v_g^2,
  \end{equation}
  % 
  where $v_g = \frac{Q_w}{S_g}$ is the water velocity.
  Here, $Q_w$ represents the instantaneous water discharge, rather than $Q_w^*$, which is used to evaluate the channel geometry.
  With this formulation, the processes in an R-channel, through which water flows subglacially, can be represented algebraically \citep{rothlisberger1972,delaney2019}.

  For purposes discussed below, width integrated shear stress is represented as $\tau_{tg}=w_{cg}\,\tau_g $.

  \subsection{Subaerial channel  parameterization}
  \label{sect:fluv}

 To parameterize the shear stress of water flowing across sediments in the subaerial channel,  the hydraulics parameterization presented in \citet{tucker1997} is implemented.
  Here, it is assumed that there is conservation of mass and a sufficiently wide channel so that the hydraulic radius is consistent with flow depth, uniform flow, and the Darcy-Weisbach relationship.
  The shear stress $\tau_f$ at the river bed is represented as
  \begin{linenomath*}
    \begin{equation}
      \label{eq:DW_tau}
      \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_p^{\frac{1}{3}}}{2}\, \Big(\frac{Q_w}{w_{cf}} \Big)^{\frac{2}{3}} \,\frac{\partial z_c}{\partial x}^{\frac{2}{3}},
    \end{equation}
  \end{linenomath*}
  where $\frac{\partial z_c}{\partial x}$ is the channel slope, and $f_p$ is the friction factor for subaerial systems.
  Channel width $w_{cf}$ is 
  \begin{equation}
    \label{eq:wcf}
    w_{cf} = k \, Q_w^e,
  \end{equation}
  % 
  where $k$ is a constant and $e$ is an exponent commonly equal to $\frac{1}{2}$ \citep{leopold1953}.
  Cross-sectional area, $ S_f$, is 
  \begin{equation}
    \label{eq:Sf}
    S_f = \frac{Q_w}{v_f},
  \end{equation}
  %
  where $v_f$ is water velocity, given by rearranging Equation \ref{eq:tau} as
 \begin{equation}
    \label{eq:vf}
    v_f = \sqrt{\frac{8\,\tau_f}{f_p\,\rho_w}}.
  \end{equation}
  %
  As in Section~\ref{sect:sub_mode}, the width integrated shear stress is $\tau_{tf}=w_{cf}\,\tau_f$.
  Hydraulic diameter of the channel is given by $D_{hf} = \frac{4\,w_{cf}\,d}{2\,d+w_{cf}}$, with flow depth $d$ determined from knowledge of $S_f$ and $w_{cf}$ in a rectangular channel.

@@

The formulation of the glacial system (Section~\ref{sect:sub_mode}) requires inputs of the hydraulic gradient (Equation \ref{eq:psi}; $\frac{\partial z_{sg}}{\partial x}$, $\frac{\partial z_{c}}{\partial x}$), water discharge, $Q_w$ representative water discharge, $Q_w^*$,  Hooke angle (Equation~\ref{eq:Hf}; $\beta$), and friction factor (Equation~\ref{eq:DW}; $f_r$). \marie{Help the reader find these quantities and equations}
  Values of $\beta$ and $f_r$ are chosen so that they produce reasonable values of water velocity ($0.6-- 1.2$\,\unit{m}\,\unit{s}$^{-1}$) in accordance with measurements made through dye-tracing experiments to examine subglacial hydraulic conditions \citep[Section~\ref{sect:sub_mode}, Figure~\ref{fig:model_outs}; e.g.][]{werder2010}.
